---
title: "<div style='text-align:center;'><span style='color:red; font-style:italic;'>Project Genre Matters: Exploring Film Audience Preferences</span></div>"
---

Welcome to our project page.

Keep an eye on this space to stay updated with my project activities.

## *1.Project Overview and Scope*

In this project, we aim to analyze how different movie genres perform in terms of audience preferences. Our primary focus is to identify genre-based patterns in film popularity and ratings, using a data-driven approach. To make the analysis more meaningful and localized, we enriched a widely-used global movie dataset by merging it with a Turkish-specific movie dataset. Through this project, we plan to discover whether certain genres are more successful among audiences and if factors like production year, runtime, or budget have a significant effect on genre-based popularity.

## *2.Data*

We based our analysis on a combined dataset. The core dataset is *ggplot2movies*, which includes information about thousands of films worldwide.\

To create a more localized perspective, we merged this dataset with a Turkish movie dataset.\

The combined data allowed us to create new variables, such as Turkish-specific ratings and popularity scores, which we will use in our comparative analysis.

### *2.1 Data Source*

-   The global movie data is sourced from the open-access *ggplot2movies* dataset, widely used in data analysis and visualization projects.

-   The Turkish-specific data was obtained from Kaggle, specifically from the [Turkish Movies Dataset](https://www.kaggle.com/datasets/emreokcular/turkish-movies) shared by Emre Okçular.

-   Both datasets provide rich information such as film titles, genres, production years, runtimes, IMDb ratings, and other useful attributes for our analysis.

### 2.*2 General Information About Data*

#### ***2.2.1 Data Combination Process***

To create a comprehensive and diverse dataset, we combined two different sources.\
First, we imported the original *ggplot2movies* dataset, which contains global movie information including variables such as title, year, length, budget, rating, votes, and genre indicators.\
Second, we processed a Turkish movie dataset, cleaned it, and expanded its structure to match the global dataset's format.

We standardized column names, added missing fields where necessary (e.g., budget, votes, genre dummies), and included a `country` variable to distinguish between Global and Turkish movies.\
After ensuring consistency across both datasets, we merged them into a single, unified dataset named `combined_movies`.\
This final dataset preserves authentic movie titles and metadata while also incorporating localized Turkish film information, making it suitable for international and regional analyses.

The combined dataset was saved in both `.RData` and `.csv` formats for flexibility in further exploration and analysis.

```{r}

# 1. Global movies verisini yükle
load("movies.RData")

# 2. Turkish movies verisini oku
turkish_movies <- read.csv("final_dataset.csv", fileEncoding = "UTF-8")

# 3. Turkish movies verisini genişlet
turkish_movies_expanded <- data.frame(
  title = turkish_movies$localized.title,
  year = as.integer(turkish_movies$runtimes),
  length = as.integer(turkish_movies$runtimes),
  budget = NA,
  rating = turkish_movies$rating,
  votes = NA,
  r1 = NA, r2 = NA, r3 = NA, r4 = NA, r5 = NA, 
  r6 = NA, r7 = NA, r8 = NA, r9 = NA, r10 = NA,
  mpaa = NA,
  Action = NA, Animation = NA, Comedy = NA, Drama = NA,
  Documentary = NA, Romance = NA, Short = NA,
  country = "Turkey"
)

# 4. Global movies'a country kolonu ekle
movies$country <- "Global"

# 5. Kolon sıralarını eşitle
turkish_movies_expanded <- turkish_movies_expanded[, names(movies)]

# 6. İkisini birleştir
combined_movies <- rbind(movies, turkish_movies_expanded)

# 7. Kaydet
save(combined_movies, file = "combined_movies.RData")

# 8. CSV de oluştur
write.csv(combined_movies, file = "combined_movies.csv", row.names = FALSE)

```

#### ***2.2.2 Data Combination Process***

The final dataset combines global and Turkish movies, resulting in a comprehensive collection of approximately 8,000 films.\

It includes a wide range of information for each movie, covering basic metadata, audience ratings, and genre classifications.

The following code blocks provide an overview of the combined movies dataset, including its structure, dimensions, variable names, sample records, and distribution patterns.

```{r}

# Load the combined dataset
load("combined_movies.RData")

# Show structure
str(combined_movies)
```

```{r}
# Number of rows and columns
dim(combined_movies)
```

```{r}

# Names of variables
names(combined_movies)

```

```{r}
# First 6 rows of data
head(combined_movies)

```

# Top 10 highest rated movies

library(dplyr)

combined_movies %\>% arrange(desc(rating)) %\>% select(title, rating, country) %\>% head(10)

```{r}

library(dplyr)
library(ggplot2)

# Top 10 Highest Rated Movies tablosu
top10_movies <- combined_movies %>%
  arrange(desc(rating)) %>%
  select(title, rating, country) %>%
  head(10)

top10_movies

# Top 10 Movies Rating Bar Chart
ggplot(top10_movies, aes(x = reorder(title, rating), y = rating, fill = country)) +
  geom_bar(stat = "identity") +
  coord_flip() +  # Başlıkları yatay gösterir
  labs(title = "Top 10 Highest Rated Movies",
       x = "Movie Title",
       y = "Rating") +
  theme_minimal()


```

```{r}

combined_movies %>%
  group_by(country) %>%
  summarize(
    Average_Rating = round(mean(rating, na.rm = TRUE), 2),
    Average_Length = round(mean(length, na.rm = TRUE), 1)
  )

```

### 2.*3 Reasons of Choice*

We selected this dataset because it offers a rich and diverse set of information about movies across different genres and periods.\

The global *ggplot2movies* dataset provides a well-organized structure for large-scale analysis, while the Turkish Movies dataset adds a localized dimension to the study, making it more relevant to regional audience preferences.\

The combination of both datasets allows us to explore general trends as well as cultural differences in movie genre popularity.\

Moreover, the variety of available variables such as genre, rating, year, runtime, and awards creates multiple opportunities for deeper analytical approaches.

### *2.4 Preprocessing*

Before starting the analysis, we needed to make sure that our dataset was clean, consistent, and ready for use.\
The original *ggplot2movies* dataset and the Turkish Movies dataset were first reviewed separately.\
We noticed that there were slight differences in the available columns, especially in genre categorization and runtime fields.

In order to combine the two datasets properly, we carried out the following preprocessing steps:

-   Selected the relevant columns from both datasets (e.g., title, year, runtime, genre, rating, votes, and award-related fields).

-   Standardized column names and data formats to ensure consistency.

-   Merged the datasets into a single file by aligning the columns.

-   Replaced Turkish-specific characters (such as "ş", "ğ", "ü", "ö", "ç", "ı") with their English equivalents to prevent encoding problems.

-   Filled missing values in fields like runtime and budget using reasonable defaults where needed.

-   Added a new `country` column to differentiate between Global and Turkish movies.

-   At the end of the preprocessing phase, we saved the final version of the dataset as an `.RData` file, making it ready for the subsequent stages of analysis.

```{r}

# 1. Import CSV Dataset
combined_movies <- read.csv("combined_movies_final.csv", fileEncoding = "UTF-8")

# 2. Save as RData
save(combined_movies, file = "combined_movies.RData")

# 3. Load the dataset
load("combined_movies.RData")
```
